Viscous Transport in Eroding Porous Media

### 2019-09-24 docs/transport/erosion_transport.tex
still editing ...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{sec:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{left off here}
In line with our previous work~\cite{qua-moo2018}, we use two meshes to
simulate erosion. The integral equation is solved by discretizing the
boundary of the geometry at a set of points distributed equally in
arclength, and then applying a collocation method
(Section~\ref{sec:spatialDiscretization}).  Depending on the proximity
of the target point to the source points, the quadrature rule is either
the trapezoid rule or the Barycentric quadrature rule
(Section~\ref{sec:bary}).  The selection algorithm that determines which
quadrature method is applied is described in Section~\ref{sec:fmm}.
Once the shear stress is computed, the bodies are eroded a single time
step by using a {\thL} discretization~\cite{hou-low-she1994}.  By using
a {\thL} formulation, the stiff curvature term is linear, treated
implicitly, and the discretization points remain equispaced in
arclength.  Finally, time stepping methods for the eroding bodies and
passive particles are described in Section~\ref{sec:time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using an integral equation formulation, we only have to discretize
the one-dimensional boundary of the domain.  Therefore, we easily
resolve complex geometries, and differentiation is performed with
spectral accuracy in Fourier space.  We discretize each eroding body
$\gamma_k$ with $N_\iin$ points and discretize the outer wall $\Gamma$
with $N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply mild smoothing
strategies~\cite{qua-moo2018} to avoid the corners that inevitably
develop in erosion.

Given the discretization points of $\bd\Omega$, applying the trapezoid
rule results in the collocation method for~\eqref{eqn:BIE} 
\begin{subequations}
\label{eqn:trapLinearSystem}
  \begin{alignat}{3}
  \UU_\ell &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^0_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^0_\ell,\yy^k_j) \eeta^k_j +
  \sum_{j=1}^{N_\out} w^0_j N_0(\yy^0_\ell,\yy^0_j)\eeta_j^0 
    \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^0_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^0_\ell),
    &&\hspace{-60pt} \ell = 1,\ldots,N_\out, \\
%
  \mathbf{0} &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^m_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^m_\ell,\yy^k_j) \eeta^k_j \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^m_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^m_\ell),
    &&\hspace{-60pt}\ell = 1,\ldots,N_\iin, \: m=1,\ldots,M, \\
%
  \llambda_m &= \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m \eeta^m_j,
  \quad \xi_m = \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m
    (\yy^m_j - \cc_m)^\perp \cdot \eeta^m_j,
    &&\hspace{-60pt} m=1,\ldots,M,
\end{alignat}
\end{subequations}
where $w^k_j$ are quadrature weights that depend on $N_\iin$,
$N_\out$, and the lengths of $\gamma^k$, and $D$ is the kernel of the
Stokes double-layer potential defined in
equation~\eqref{eqn:velocityDLP}.  The linear
system~\eqref{eqn:trapLinearSystem} is a well-conditioned second-kind
integral equation that is solved iteratively with GMRES.  If the number
of discretization points is sufficiently large, then the solution
of~\eqref{eqn:trapLinearSystem} is an accurate approximation of the
density function, Stokeslets, and rotlets.   Then, if $\xx \in \Omega$
is sufficiently far from $\bd\Omega$, an approximation of the velocity,
deformation tensor, and vorticity due to the double-layer potential is
found by applying the trapezoid rule to \eqref{eqn:velocityDLP},
\eqref{eqn:deformationDLP}, and~\eqref{eqn:vorticityDLP}, respectively.
In particular,
\begin{align}
  \label{eqn:trap}
  \uu(\xx) &= \sum_{j=1}^{N_\out} w^0_j D(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D(\xx,\yy^k_j) \eeta^k_j, \\
%
  \ssigma(\xx) &= \sum_{j=1}^{N_\out} w^0_j \mathbf{D}_\ssigma(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j \mathbf{D}_\ssigma(\xx,\yy^k_j)
  \eeta^k_j, \\
%
  w(\xx) &= \sum_{j=1}^{N_\out} w^0_j D_\omega(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D_\omega(\xx,\yy^k_j)
  \eeta^k_j.
\end{align}
The contributions due to the Stokeslets and rotlets require no
quadrature and are easily included in the velocity, deformation tensor,
and vorticity.  Finally, to compute the shear stress, we apply the
trapezoid rule with Fourier differentiation to approximate the jump
term~\eqref{eqn:deformationJump} of the deformation tensor, and then the
tensor is applied to the normal and tangent vectors as defined in
equation~\eqref{eqn:shearStress}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature Formulas}
\label{sec:bary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since the error of the trapezoid rule applied to periodic functions
depends on the derivative of the integrand~\cite{tre-wei2014}, the
trapezoid rule applied in Section~\ref{sec:spatialDiscretization} is
only appropriate if the target point is sufficiently separated from all
eroding bodies and the outer wall.  Therefore, if the trapezoid rule is
applied when bodies are in near-contact, or when a layer potential is
evaluated at a point close to $\bd\Omega$, then the results become
unreliable and the method ultimately becomes unstable.  Therefore, we
require a quadrature method whose error is bounded independent of the
target location.

We showed in Sections~\ref{sec:DLPcomplex} and~\ref{sec:gradDLPcomplex}
that the velocity, shear stress, and vorticity of the double-layer
representation can all be written in terms of Cauchy integrals and its
first two derivatives.  Therefore, if we develop quadrature rules with a
uniformly bounded error for Cauchy integrals and its
derivatives~\eqref{eqn:cauchy}, then the quadrature method can be used
to accurately compute the velocity, shear stress, and vorticity.
Ioakimidis et al.~\cite{ioa-pap-per1991} developed quadrature rules,
that we call {\em Barycentric quadrature rules}, to compute Cauchy
integrals and their derivatives with an error bound that is independent
of $x$.  Then, Barnett et al.~\cite{bar-wu-vee2015} applied these
quadrature methods to compute the Stokes double-layer potential
representation of the velocity~\eqref{eqn:velocityDLP}.  We build on
this method so that the shear stress and vorticity can be computed with
a uniform error bound.  We briefly summarize the Baryncentric quadrature
method for computing the velocity, and we develop similar quadrature
rules for the gradient of the Stokes double-layer potential.

@@

Again, given a complex-valued density function $\eta$, the boundary data
of $v[\eta](x)$, as defined in equation~\eqref{eqn:laplaceComplex},
satisfies the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.
Since the boundary data of $v$ differs when considering $x \in \Omega$
and $x \in \Omega^c$, we denote the boundary data as $v^-$ for $x \in
\Omega$, and as $v^+$ for $x \in \Omega^c$.  Rather than directly
applying the trapezoid rule to $v(x)$, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
    \quad x \in \Omega.
\end{align}
Since the integrand is bounded as $x$ approaches $\bd\Omega$, we can
apply the trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the interior Barycentric quadrature rule
\begin{align}
  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}, \quad x \in \Omega.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the exterior Barycentric quadrature rule is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvExterior}
\end{align}

@@

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of Ioakimidis et al.~\cite{ioa-pap-per1991} (see equation (2.12)).
We start with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
Similar to the quadrature rule for $v'(x)$ requiring the value $v(x)$,
the quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$.
Both these quantities are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form stable eroding simulations of
nearly-touching grains, and to study dynamics of the flow in regions
arbitrarily close to eroding grains.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{sec:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  By using a fast
summation method, such as the fast multipole method~\cite{gre-rok1987},
this cost can be reduced to $\mathcal{O}(N)$ operations.  However, the
Barycentric quadrature rules require several $N$-body calculations, and
they must be done for each component of $\bd\Omega$ individually.  The
resulting cost would be prohibitive, so we use a hybrid method that
combines the Barycentric quadrature rules and the trapezoid rule.  In
short, the fast multipole accelerated trapezoid rule is used to compute
the velocity at all the discretization points, and the non-accelerated
Barycentric quadrature rule is used to locally correct the errors
introduced by the trapezoid rule.  Note that the source points of the
layer potential is always one of the eroding bodies or the outer wall,
but the target point can either be on another component of $\bd\Omega$
or it can be in the fluid bulk $\Omega$.

To compute the velocity double-layer potential, we start by applying the
trapezoid rule~\eqref{eqn:trap} accelerated with the FMM.  This
calculation requires $\mathcal{O}(N)$ operations, and we call the
resulting velocity $\vv_\trap(\xx)$.  Depending on the number of
discretization points $N_\iin$ and $N_\out$, there is a region separated
from $\bd\Omega$ where the error of $\vv_\trap(\xx)$ is less than a
user-specified tolerance.  However, for the other points in $\Omega$,
the trapezoid rule needs to be replaced with a more accurate quadrature
rule.  Note, however, that since a point is typically only close to one
or two components of $\bd\Omega$, only the contribution of these nearby
bodies need to be replaced.  Assuming that $\xx$ is too close to
$\gamma_k$, we first subtract the inaccurate approximation of the
double-layer potential due to $\gamma_k$.  Then, the Barycentric
quadrature rule is used to compute the velocity due to $\gamma_k$ with
more accuracy.  That is, the final velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
  \label{eqn:velocityDecomp}
\end{align}
where $\vv^k_\bary(\xx)$ is the velocity at $\xx$ resulting from
applying the Barycentric quadrature rule to the double-layer potential
due to $\gamma_k$.  This strategy naturally extends to points that are
close to $\Gamma$, and to the scenario when $\xx$ is simultaneously
close to multiple components of $\bd\Omega$.  While the term
$\vv_\trap(\xx)$ in equation~\eqref{eqn:velocityDecomp} is computed for
all target points using the FMM, the other two terms are computed with a
direct summation.  However, these terms are only required for target
points near an eroding body or the outer wall, and these points make up
only a small fraction of the total number of points.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{sec:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In line with our previous work~\cite{qua-moo2018}, we use two meshes to
simulate erosion. The integral equation is solved by discretizing the
boundary of the geometry at a set of collocation points distributed
equally in arclength (Section~\ref{sec:spatialDiscretization}).
Depending on the proximity of the target point to the source points, the
quadrature rule is either the trapezoid rule or the Barycentric
quadrature rule (Section~\ref{sec:bary}).  The selection algorithm that
determines which quadrature method is applied is described in
Section~\ref{sec:fmm}.  Once the shear stress is computed, the bodies
are eroded a single time step by using a {\thL}
discretization~\cite{hou-low-she1994}.  The time stepping methods for
erosion and passive particles are described in Section~\ref{sec:time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we use a BIE formulation, we only have to discretize the
one-dimensional boundary of the domain.  Therefore, we easily resolve
complex geometries, and differentiation is performed with spectral
accuracy in Fourier space.  We discretize each eroding body $\gamma_k$
with $N_\iin$ points and discretize the outer wall $\Gamma$ with
$N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply mild smoothing
strategies~\cite{qua-moo2018} to avoid the corners that inevitably
develop in erosion.

Given the discretization points of $\bd\Omega$, applying the trapezoid
rule results in the collocation method for~\eqref{eqn:BIE} 
\begin{subequations}
\label{eqn:trapLinearSystem}
  \begin{alignat}{3}
  \UU_\ell &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^0_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^0_\ell,\yy^k_j) \eeta^k_j +
  \sum_{j=1}^{N_\out} w^0_j N_0(\yy^0_\ell,\yy^0_j)\eeta_j^0 
    \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^0_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^0_\ell),
    &&\hspace{-60pt} \ell = 1,\ldots,N_\out, \\
%
  \mathbf{0} &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^m_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^m_\ell,\yy^k_j) \eeta^k_j \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^m_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^m_\ell),
    &&\hspace{-60pt}\ell = 1,\ldots,N_\iin, \: m=1,\ldots,M, \\
%
  \llambda_m &= \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m \eeta^m_j,
  \quad \xi_m = \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m
    (\yy^m_j - \cc_m)^\perp \cdot \eeta^m_j,
    &&\hspace{-60pt} m=1,\ldots,M,
\end{alignat}
\end{subequations}
where $w^k_j$ are quadrature weights that depend on $N_\iin$, $N_\out$,
and the lengths of $\gamma^k$ and $\Gamma$, and $D$ is the kernel of the
Stokes double-layer potential defined in
equation~\eqref{eqn:velocityDLP}.  The linear
system~\eqref{eqn:trapLinearSystem} is a well-conditioned second-kind
integral equation that is solved iteratively with GMRES.  If the number
of discretization points is sufficiently large, then the solution
of~\eqref{eqn:trapLinearSystem} is an accurate approximation of the
density function, Stokeslets, and rotlets.   Then, if $\xx \in \Omega$
is sufficiently far from $\bd\Omega$, an approximation of the velocity,
deformation tensor, and vorticity due to the double-layer potential is
found by applying the trapezoid rule to \eqref{eqn:velocityDLP},
\eqref{eqn:deformationDLP}, and~\eqref{eqn:vorticityDLP}, respectively.
In particular,
\begin{align}
  \label{eqn:trap}
  \uu(\xx) &= \sum_{j=1}^{N_\out} w^0_j D(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D(\xx,\yy^k_j) \eeta^k_j, \\
%
  \ssigma(\xx) &= \sum_{j=1}^{N_\out} w^0_j \mathbf{D}_\ssigma(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j \mathbf{D}_\ssigma(\xx,\yy^k_j)
  \eeta^k_j, \\
%
  w(\xx) &= \sum_{j=1}^{N_\out} w^0_j D_\omega(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D_\omega(\xx,\yy^k_j)
  \eeta^k_j.
\end{align}
The contributions due to the Stokeslets and rotlets require no
quadrature and are easily included in the velocity, deformation tensor,
and vorticity.  Finally, to compute the shear stress, we apply the
trapezoid rule with Fourier differentiation to approximate the jump
term~\eqref{eqn:deformationJump} of the deformation tensor, and then the
tensor is applied to the normal and tangent vectors as defined in
equation~\eqref{eqn:shearStress}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature Formulas}
\label{sec:bary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we are applying quadrature to periodic smooth functions, the
trapezoid rule is spectrally accurate~\cite{tre-wei2014}. However, since
the derivative of the integrands grows as the target point $\xx$
approaches $\bd\Omega$, the error bound of the trapezoid rule depends on
the location of the target point. Therefore, if the trapezoid rule, as
described in Section~\ref{sec:spatialDiscretization}, is applied when
bodies are in near-contact, or when a layer potential is evaluated at a
point close to $\bd\Omega$, then the results become unreliable and the
method ultimately becomes unstable.  Thus, we require a quadrature
method with an error bound that is independent of the target location.

We showed in Sections~\ref{sec:DLPcomplex} and~\ref{sec:gradDLPcomplex}
that the velocity, shear stress, and vorticity of the double-layer
representation can all be written as the sum of Cauchy integrals and its
first two derivatives.  Therefore, we can accurately compute the
velocity, shear stress, and vorticity at any target location if we
develop a quadrature rules with a uniform error bounded for Cauchy
integrals and its derivatives~\eqref{eqn:cauchy}.  Ioakimidis et
al.~\cite{ioa-pap-per1991} developed quadrature rules, that we call {\em
Barycentric quadrature rules}, to compute Cauchy integrals and their
derivatives with an error bound that is independent of $x \in \CC$.
Then, Barnett et al.~\cite{bar-wu-vee2015} used these quadrature rules
to compute the Stokes double-layer potential representation of the
velocity~\eqref{eqn:velocityDLP}, and we briefly summarize their method.
Then, we build on this work to accurately compute the gradient of the
Stokes double-layer potential so that we form the shear stress and
vorticity.

@@

Again, given a complex-valued density function $\eta$, the boundary data
of $v[\eta](x)$, as defined in equation~\eqref{eqn:laplaceComplex},
satisfies the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.
Since the limiting boundary data of $v$ differs when considering $x \in
\Omega$ and $x \in \Omega^c$, we denote the boundary data as $v^-$ for
$x \in \Omega$, and as $v^+$ for $x \in \Omega^c$.  Rather than directly
applying the trapezoid rule to approximate $v(x)$ in
equation~\eqref{eqn:cauchy}, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
    \quad x \in \Omega.
\end{align}
Since the integrand is bounded for all $x \in \Omega$, we can apply the
trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the interior Barycentric quadrature rule
\begin{align}
  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}, \quad x \in \Omega.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the exterior Barycentric quadrature rule is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvExterior}
\end{align}

@@

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of Ioakimidis et al.~\cite{ioa-pap-per1991} (see equation (2.12)).
We start with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$,
and theses are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form stable eroding simulations of
nearly-touching grains, and to study dynamics of the flow in regions
arbitrarily close to eroding grains.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{sec:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using the Barycentric quadrature rule, the velocity, shear stress,
and vorticity are computed with an error that is bounded independent of
the target location. However, applied directly, it requires
$\mathcal{O}(N^2)$ operations.  By using a fast summation method, such
as the fast multipole method~\cite{gre-rok1987}, this cost can be
reduced to $\mathcal{O}(N)$ operations.  However, the Barycentric
quadrature rules require several $N$-body calculations, and they must be
done for each component of $\bd\Omega$ individually.  The resulting cost
would be prohibitive, so we introduce a hybrid method that combines the
Barycentric quadrature rule and the trapezoid rule.  In short, the fast
multipole accelerated trapezoid rule is used to compute the velocity at
all the discretization points, and the non-accelerated Barycentric
quadrature rule is used to locally correct the errors introduced by the
trapezoid rule.  Note that the source points of the layer potential is
always one of the eroding bodies or the outer wall, but the target point
can either be on another component of $\bd\Omega$ or it can be in the
fluid bulk $\Omega$.

\todo[inline]{left off here}
To compute the velocity double-layer potential, we start by applying the
trapezoid rule~\eqref{eqn:trap} accelerated with the FMM.  This
calculation requires $\mathcal{O}(N)$ operations, and we call the
resulting velocity $\vv_\trap(\xx)$.  Depending on the number of
discretization points $N_\iin$ and $N_\out$, there is a region separated
from $\bd\Omega$ where the error of $\vv_\trap(\xx)$ is less than a
user-specified tolerance.  However, for the other points in $\Omega$,
the trapezoid rule needs to be replaced with a more accurate quadrature
rule.  Note, however, that since a point is typically only close to one
or two components of $\bd\Omega$, only the contribution of these nearby
bodies need to be replaced.  Assuming that $\xx$ is too close to
$\gamma_k$, we first subtract the inaccurate approximation of the
double-layer potential due to $\gamma_k$.  Then, the Barycentric
quadrature rule is used to compute the velocity due to $\gamma_k$ with
more accuracy.  That is, the final velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
  \label{eqn:velocityDecomp}
\end{align}
where $\vv^k_\bary(\xx)$ is the velocity at $\xx$ resulting from
applying the Barycentric quadrature rule to the double-layer potential
due to $\gamma_k$.  This strategy naturally extends to points that are
close to $\Gamma$, and to the scenario when $\xx$ is simultaneously
close to multiple components of $\bd\Omega$.  While the term
$\vv_\trap(\xx)$ in equation~\eqref{eqn:velocityDecomp} is computed for
all target points using the FMM, the other two terms are computed with a
direct summation.  However, these terms are only required for target
points near an eroding body or the outer wall, and these points make up
only a small fraction of the total number of points.
